---
title: "Mosaics: A. cervicornis and ticket community"
author: "Stephanie Martinez"
date: "10/3/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Benthic community and health factors of the mosaics

```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)
library(artyfarty)
library(patchwork)
library(tidyr)
library(tibble)
library(magrittr)
library(grafify)
library(vegan) #PERMNANOVA and SIMPER
library(purrr) #map function
library(gt)

```

### Data

```{r}
cover <- read.csv("data/cover_groups.csv", sep = ";",dec = ".")  %>% 
  slice(.,-(9:10)) %>%  #remove BQV (onlt 2 transects)
  mutate(site = factor(site,levels = c("HDMS","IF","CAN", "IL", "ESP","NOR", "VEN", "RAB", "MAD", "BM1", "BM2")))  #mutate to reorder locations to match map coordinates
  
  
wbd <- read.csv("data/wbd.csv", sep = ";", dec = ".") %>% 
  mutate(site = factor(site,levels = c("HDMS","IF","CAN", "IL", "ESP","NOR", "VEN", "RAB", "MAD", "BM1", "BM2"))) #mutate to reorder locations to match map coordinates
```

## Benthic community

### Stacked barplot of benthic community

```{r}
# Merging dead coral categories  and deleting minor categories 

cover2 <- cover %>% 
  mutate(dead_ac=rdead_ac+odead_ac) %>% 
  select(-rdead_ac, -odead_ac, - other, -cyano, -millepora)

#Reshaping data frame for stacked plot

cover_mean <-  aggregate(cover2[, 4:13], list(cover2$site),mean) %>% 
  pivot_longer(scleractinian:dead_ac, names_to = "group",values_to = "cover") %>% 
  rename(site=Group.1) 

# Invert abiotic groups so they are below y=0

cover_mean2<- cover_mean %>%
  mutate(coverInv = ifelse(group == "sand" | group == "dead_ac" | group == "rubble", -1*cover, cover))

cover_mean2$group <- as.factor(cover_mean2$group) #to be able to order groups in each bar
 
# Plot

p1 <- ggplot(cover_mean2) +
  geom_bar(position="stack", stat="identity", 
           aes(fill=factor(group,levels=c("sponge", "octocoral","halimeda", 
                                          "turf.algae", "algae",
                                          "scleractinian", "cervicornis", 
                                          "dead_ac", "rubble", "sand")), #Group order inside each bar
               y=coverInv, x=site)) +
  scale_fill_manual(values = c("#AA4499","#882255","#CC6677","#117733",
                               "#44AA99","#0077BB","#332288",
                               "#26272C","#4C4E57","#A8AAB3"),
                    labels = c("Sponge", "Octocoral","Halimeda",
                               "Turf algae","Macroalgae","Scleractinian",
                               "A. cervicornis", "Dead coral","Rubble","Sand")) +
  theme_scientific()+
  theme(legend.title = element_blank()) +
  labs(x="", y="Cover (%)")+
  scale_y_continuous(breaks = pretty(cover_mean2$coverInv),
                     labels = abs(pretty(cover_mean2$coverInv)))
  
p1

ggsave(p1, filename = "figures/thicket_comm.pdf")

```

## *A. cervicornis* live cover, recent and dead mortality comparison

Formatting data frames for barplots

```{r}

cervi <- cover %>% 
  select(sector,site, transect, cervicornis, rdead_ac, odead_ac)

cervi_mean <- aggregate(cervi[, 4:6], list(cervi$site),mean) %>% 
  pivot_longer(c(cervicornis, rdead_ac, odead_ac),names_to = "group",values_to = "mean") 
cervi_sd <- aggregate(cervi[, 4:6], list(cervi$site),sd) %>% 
  pivot_longer(c(cervicornis, rdead_ac, odead_ac),names_to = "group",values_to = "sd") %>% 
  select(-Group.1, -group)
  
cervi2 <- bind_cols(cervi_mean, cervi_sd) %>% #merging mean and sd in one df
  rename(site = Group.1) %>% 
  slice(-(7:9)) %>% 
  mutate(site = factor(site,levels = c("HDMS","IF","CAN", "IL", "ESP","NOR", "VEN", "RAB", "MAD", "BM1", "BM2"))) #mutate to reorder locations to match map coordinates
```

Barplot: A. cervicornis live cover, recent and old mortality

```{r}

p2 <- ggplot(cervi2, aes(x=site, y=mean, fill=group)) + 
  geom_bar(stat="identity", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean, ymax=mean+sd), width=.2,
                 position=position_dodge(.9)) +
  theme_scientific()+
  scale_fill_grafify(palette = "muted",
                    labels = c("Live cover", "Old mortality", "Recent mortality")) +
  theme(legend.title = element_blank(), legend.position="bottom",plot.title = element_text(hjust = 0)) +
  labs(x="", y="Relative cover (%)", title="a)")
p2

#White band disease boxplot

p3 <- ggboxplot(wbd, x = "site", y = "freq", xlab="", ylab="WBD relative frequency (%)", title = "b)")+
  theme_scientific()+
  theme(plot.title = element_text(hjust = 0))
p3  

#Composing plots

p4 <- p2/p3
p4

ggsave(p4, filename = "figures/thicket_status.pdf")
```

## Statistics

### Community PERMANOVA

To determine if there were differences between factor levels (Sector: Fixed; Site: random, nested to sector) a PERMANOVA using Bray-Curtis similarity matrix was done.

```{r}
# PERMANOVA  
cover_permanova <- adonis(cover[,c(4:17)]~sector+site%in%sector,
                    method = "bray",data=cover)

cover_permanova


# SIMPER per sector

sim_sector <- simper(cover[,c(4:17)],cover$sector)
sim_sector_sum <- summary(sim_sector,ordered=TRUE)

# SIMPER per site

sim_site <- simper(cover[,c(4:17)],cover$site)
sim_site_sum <- summary(sim_site,ordered=TRUE)

# Selecting the first 5 groups that contributed more to the dissimilarity between sites to determine which groups contributed more frequently to dissimilarities between sectors and sites. The top 5 groups were selected since in most cases these accounted for at least 70% of the dissimilarities. 

#sector
map(sim_sector_sum,function(x){
  slice(x,c(1:5)) %>% 
    rownames_to_column("group") %>% 
    select(group)
  
}) %>% 
  unlist() %>% 
  table() %>% 
  data.frame() %>% 
  rename("group"=".") %>% 
  arrange(desc(Freq)) %>% 
  mutate(Freq_p=round((Freq/3)*100,1)) #Divide freq by number of lists
  

# site
map(sim_site_sum,function(x){
  slice(x,c(1:5)) %>% 
    rownames_to_column("group") %>% 
    select(group)
  
}) %>% 
  unlist() %>% 
  table() %>% 
  data.frame() %>% 
  rename("group"=".") %>% 
  arrange(desc(Freq)) %>% 
  mutate(Freq_p=round((Freq/55)*100,1)) #Divide freq by number of lists
  

```

We found that there were differences in the community associated with the geographical sectors and sites. Geographical sectors explained 10% of the variation, while sites explained 52% of the variation

According to the SIMPER by sector groups, cervicornis live tissue, old dead mortality, rubble and sand where the groups that contributed mostly to variation across sector.

In terms of sites, the groups that contributed more to dissimilarities were: cervicornis cover, old mortality, rubble, sand. And in a lesser extent octocoral and macroalgae.


### WBD PERMANOVA

```{r}

wbd_permanova <- adonis(wbd[,4]~sector+site%in%sector,
                    method = "euclidean",data=wbd)

wbd_permanova

```

The ocurrence (frequency) of WBD was different across sites, but not across geographic sectors

## Extracting statistic table results with style

Community PERMANOVA results

```{r}

table_permanova <- as.data.frame(cover_permanova$aov.tab) %>% 
  rownames_to_column("VariationSource") %>%
  rename(pvalue = 7) %>% 
  mutate_at(vars(SumsOfSqs, MeanSqs, F.Model, R2), funs(round(., 2))) %>% 
  gt() %>% 
  fmt_missing(
    columns = 2:7,
    missing_text = ""
  ) %>% 
  cols_move_to_end(columns = R2) %>% 
  cols_label(
    VariationSource = md("**Variation source**"),
    Df = md("**df**"),
    SumsOfSqs = md("**Sum of Squares**"),
    MeanSqs = md("**Mean Squares**"),
    F.Model = md("**F-Model**"),
    pvalue = md("**p-value**"),
    R2 = md("**R2**")
  ) %>% 
  cols_align(align = "center", columns = c("Df","SumsOfSqs","MeanSqs",
                                           "F.Model","R2","pvalue")
  ) 

table_permanova

# Save table in directory

gtsave(filename = "figures/community_permanova.tex", data = table_permanova)
gtsave(filename = "figures/community_permanova.png",  data = table_permanova)


```

WBD permanova table

```{r}

table_wbd <- as.data.frame(wbd_permanova$aov.tab) %>% 
  rownames_to_column("VariationSource") %>%
  rename(pvalue = 7) %>% 
  mutate_at(vars(SumsOfSqs, MeanSqs, F.Model, R2), funs(round(., 2))) %>% 
  gt() %>% 
  fmt_missing(
    columns = 2:7,
    missing_text = ""
  ) %>% 
  cols_move_to_end(columns = R2) %>% 
  cols_label(
    VariationSource = md("**Variation source**"),
    Df = md("**df**"),
    SumsOfSqs = md("**Sum of Squares**"),
    MeanSqs = md("**Mean Squares**"),
    F.Model = md("**F-Model**"),
    pvalue = md("**p-value**"),
    R2 = md("**R2**")
  ) %>% 
  cols_align(align = "center", columns = c("Df","SumsOfSqs","MeanSqs",
                                           "F.Model","R2","pvalue")
  ) 

table_wbd

# Save table in directory

gtsave(filename = "figures/wbd_permanova.tex", data = table_wbd)
gtsave(filename = "figures/wbd_permanova.png",  data = table_wbd)


```


Uploading and formatting DISTLM table (analysis donde using Primer 6+)

```{r}

distlm <- read.csv("data/distlm.csv",sep = ";",dec = "." ) %>% 
  as.data.frame() %>% 
  gt() %>% 
  cols_label(
    Variable = md("**Variable**"),
    SS.trace. = md("**Sum Squares**"),
    Pseudo.F = md("**Pseudo-F**"),
    P = md("**p-value**"),
    Prop = md("**Prop**")
  ) %>% 
  cols_align(align = "center", columns = c("SS.trace.", "Pseudo.F","P","Prop")
  ) 

distlm

gtsave(filename = "figures/distlm.tex", data = distlm)
gtsave(filename = "figures/distlm.png",  data = distlm)

```




